Analyzing Toronto's Dining Scene: Public Food Safety and Economic Indicators in Focus

### 2024-01-25 Outputs_Folder/Paper/toronto_dinesafe_paper.qmd
Fixed typos and updated figure 1, 3 graphs

dataset <- read_csv("Inputs Folder/Data/Clean_Dinesafe_Data.csv")

@@

## Toronto Dine Safe Pass Status
The report extracted the complete data through R and classified the number of institutions by Establishment Status in the Dinesafe dataset [@TorontoDineSafe2021], specifically showing two categories: "Conditional Pass" and "Pass" as @fig-Number_of_Establishments_by_Status shown. Taken together, the count for the immediate "Pass" category is significantly higher, indicated by the higher bar on the right, indicating that more businesses are fully compliant with health inspection standards than those receiving a "Conditional Pass." "Conditional pass" indicates that certain businesses must address specific issues to fully comply with health standards. This visual demonstrates that most businesses inspected adhere to the required health and safety guidelines.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-Number_of_Establishments_by_Status
#| fig-cap: Number of Establishments by Status
library(tidyverse)

dataset <- read_csv("Inputs Folder/Data/Clean_Dinesafe_Data.csv")

status_count <- dataset %>%
  group_by(`Establishment Status`) %>%
  summarize(Count = n())

ggplot(status_count, aes(x = `Establishment Status`, y = Count)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(
       x = "Establishment Status", y = "Count")
```

@@

```
\break
## Actions from Infraction Companies
To further evaluate the actions these violating companies took after being flagged, the report used the "Minor" level with the most violations for analysis and created a pie chart to more clearly visualize the data [@fig-distribution_minor_severity_infractions]. Please note that the coding of this chart refers to the "Pie Chart customization" website [@rggpiechart]. The chart is dominated by the green proportion, representing the behaviour of "notice to comply." This suggests that in most cases of minor level, immediate closure or fines are not the first action Toronto Public Health considers; instead, businesses are allowed to make amends within a certain period.

```{r}
#| echo: false
#| message: false
#| label: fig-distribution_minor_severity_infractions
#| fig-cap: Distribution of Food Inspection Infractions by Severity
#| warning: false
library(ggplot2)
library(dplyr)
library(lubridate)

dataset <- read_csv("Inputs Folder/Data/Clean_Dinesafe_Data.csv")

@@

ggplot(minor_severity_actions, aes(x = "", y = n, fill = Action)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") + # This creates a pie chart, remove this line for a stacked bar chart
  theme_minimal() +
  labs(
       x = "Action", y = "Count")

@@

cpi_data <- read_csv("Inputs Folder/Data/Clean_CPI_Data.csv") %>%
  filter(measure_name == "CPI - Consumer Price Index - Toronto (% Change)") %>%
  rename(CPI = measure_value) %>%
  mutate(Year = as.numeric(year)) %>%
  group_by(Year) %>%
  summarize(Average_CPI = mean(CPI, na.rm = TRUE))

dataset <- read_csv("Inputs Folder/Data/Clean_Dinesafe_Data.csv") %>%
  mutate(Inspection_Date = as.Date(`Inspection Date`, format = "%Y-%m-%d")) %>%
  mutate(Year = as.numeric(format(Inspection_Date, "%Y"))) %>%
  group_by(Year) %>%
  summarize(Total_Infractions = n())

@@

file_path <- "Inputs Folder/Data/Clean_Dinesafe_Data.csv"

@@

file_path <- "Inputs Folder/Data/Clean_CPI_Data.csv"

>>>>

dataset <- read_csv("/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv")

@@

## Toronto Dine Safe Pass Status
The report extracted the complete data through R and classified the number of institutions by Establishment Status in the Dinesafe dataset [@TorontoDineSafe2021], specifically showing two categories: "Conditional Pass" and "Pass" as @fig-Number_of_Establishments_by_Status shown. Taken together, the count for the immediate "Pass" category is significantly higher with number of 76436, indicated by the higher bar on the right, indicating that more businesses are fully compliant with food safety inspection standards than those 369 businesses receiving a "Conditional Pass." "Conditional pass" indicates that certain businesses must address specific issues to fully comply with health standards. This visual demonstrates that most businesses inspected adhere to the required health and safety guidelines.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-Number_of_Establishments_by_Status
#| fig-cap: Number of Establishments by Status
library(tidyverse)
library(readr)

dataset <- read_csv("/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv")

status_count <- dataset %>%
  group_by(`Establishment Status`) %>%
  summarize(Count = n(), .groups = 'drop') # It's good practice to drop groups after summarizing

ggplot(status_count, aes(x = `Establishment Status`, y = Count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count), vjust = -0.3, size = 3.5) + # Adds text on top of the bars
  theme_minimal() +
  labs(
       x = "Establishment Status", y = "Count")

@@

```
\break
## Actions from Infraction Companies
To further evaluate the actions these violating companies took after being flagged, the report used the "Minor" level with the most violations for analysis and created a bar chart to more clearly visualize the data [@fig-distribution_minor_severity_infractions]. Please note that the coding of this chart refers to the "Chart customization" website [@rggpiechart]. The chart is dominated by the green proportion, representing the behaviour of "notice to comply." This suggests that in most cases of minor level, immediate closure or fines are not the first action Toronto Public Health considers; instead, businesses are allowed to make amends within a certain period.

```{r}
#| echo: false
#| message: false
#| label: fig-distribution_minor_severity_infractions
#| fig-cap: Distribution of Food Inspection Infractions by Severity
#| warning: false
library(tidyverse)
library(readr) # Make sure to load readr for read_csv()

dataset <- read_csv("/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv")

@@

ggplot(minor_severity_actions, aes(x = Action, y = n, fill = Action)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(
       x = "Action", y = "Count") +
  coord_flip() # Flip the coordinates to make horizontal bars

@@

cpi_data <- read_csv("/cloud/project/Inputs Folder/Data/Clean_CPI_Data.csv") %>%
  filter(measure_name == "CPI - Consumer Price Index - Toronto (% Change)") %>%
  rename(CPI = measure_value) %>%
  mutate(Year = as.numeric(year)) %>%
  group_by(Year) %>%
  summarize(Average_CPI = mean(CPI, na.rm = TRUE))

dataset <- read_csv("/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv") %>%
  mutate(Inspection_Date = as.Date(`Inspection Date`, format = "%Y-%m-%d")) %>%
  mutate(Year = as.numeric(format(Inspection_Date, "%Y"))) %>%
  group_by(Year) %>%
  summarize(Total_Infractions = n())

@@

file_path <- "/cloud/project/Inputs Folder/Data/Clean_Dinesafe_Data.csv"

@@

file_path <- "/cloud/project/Inputs Folder/Data/Clean_CPI_Data.csv"
